---
title: "Project 1 - API Vignette"
author: "Tommy King"
date: "`r Sys.Date()`"
output:
  github_document:
    html_preview: false
    toc: true
    toc_dept: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path = "Images/")
```  
  
# Introduction  
In this vignette, we're going to be taking a look at the PokeAPI, writing some functions to interact with it, and doing some exploratory data analysis on the information we're grabbing.  
  
# Requirements  
For the work done here, we will be using the `tidyverse` package for all of our data manipulation, the `httr` package for making https requests to the API and `jsonlite` for interacting with json files. The code chunk for loading in these packages is included below  
  
```{r packages, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(httr)
library(jsonlite)
```  
  
# Functions Written for (Gotta) Catching All of the Data  
  
## Damage Type Analysis Function
In this section of functions, we will make some tools for querying the Pokemon type data so we can do some analysis on the balance between various types and their damage effectiveness vs. other types. We'll write one master function here that takes the user argument of a type and a specification of what kind of damage class we're interested in:  
 * `super_effective` (double damage **from** the given type),  
 * `not_very` (half damage **from** the given type)  
 * `no_effect` (no damage **from** the given type)  
 * `super_against` (double damage **against** the given type)  
 * `not_very_against` (half damage **against** the given type)  
 * `no_effect_against` (no damage **against** the given type)  
  
Our function below will query the `type` endpoint on the PokeAPI. It's going to grab the JSON at the url we build using the type we take in as an argument, and then drills down onto the desired damage class based on the other argument we're taking in. The return value here is a vector of the names of the other type(s) that fit the specifications.  
  
```{r damage, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
damage <- function(type, damage_class){
  type_info <- fromJSON(paste0("https://pokeapi.co/api/v2/type/",type,"/"))
 
  attrib <- switch(damage_class, "super_effective" = "double_damage_to",
                                 "not_very" = "half_damage_to",
                                 "no_effect" = "no_damage_to",
                                 "super_against" = "double_damage_against",
                                 "not_very_against" = "half_damage_against",
                                 "no_effect_against" = "no_damage_against")
  
  damage_list <- type_info$damage_relations[[attrib]]
  return(damage_list$name)
}

# What Pokemon types are ground types super effective against?
damage("ground", "super_effective")
```
  
## Matchup Matrix Generation.  
Next, we're going to write another function that will help us aggregate some of this data based on a "good" or "bad" damage trait for a given type. "Good" outcomes are when the given type is super effective against another type, or when another type is not very effective, or does no damage to it. "Bad" outcomes are the inverse, when another type is super effective against the given type, or when the given type is not very effective or does no damage to another type. To give this function a little more utility, we're going to give the user the opportunity to weight these outcomes (`extreme_weight` corresponds to the good outcome of being super effective and multiplied by -1 corresponds to the bad outcomes of doing no damage and having an opposing type be super effective. `modest_weight` corresponds to the good outcomes of having an opposing type do half or no damage and multiplied by -1 corresponds to the bad outcome of doing half damage against another type.) We then create a matrix of all the possible type matchups and fill in a value based on the weights and the data we pull from the previous function.  

We'll make a vector of all the possible types, use this vector for both the rows and columns of our matrix and then fill in the individual cells by iterating through all the types and getting the corresponding damage lists from our previous function. Note that we're adding the weight values here to the matchup matrix for each new damage class we find in the hopes of estimating the total advantage or disadvantage that the type in the row has versus the type in the column.  
  
```{r matchups, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, cache=TRUE}
generate_matchups <- function(extreme_weight = 2, modest_weight = 1){  
  types <- c("normal", "fire", "water", "electric", "grass", "ice", "fighting", "poison", "ground",          "flying", "psychic", "bug", "rock", "ghost", "dragon", "dark", "steel", "fairy")
  
  matchups <- matrix(0, 18, 18)
  rownames(matchups) <- as.list(types)
  colnames(matchups) <- as.list(types)
  
  for (t in types)
  {
    super <- damage(t, "super_effective")
    half <- damage(t, "not_very")
    no <- damage(t, "no_effect")
    
    for (s in super)
    {
      matchups[t,s] = matchups[t,s] + extreme_weight
      matchups[s,t] = matchups[s,t] + (-1 * extreme_weight)
    }
    for (h in half)
    {
      matchups[t,h] = matchups[t,h] + (-1 * modest_weight)
      matchups[h,t] = matchups[h,t] + modest_weight
    }
    for (n in no)
    {
      matchups[t,n] = matchups[t,n] + (-1 * extreme_weight)
      matchups[n,t] =matchups[n,t] + extreme_weight
    }
  }
  
  return(matchups)
}

# Taking a look at our generated matrix with the default weights
matchups <- generate_matchups()
matchups
```
  
## Pokemon Type Distribution Function  
Finally, we're going to make a function that will generate the distribution of the number of Pokemon belonging to each type so that we can factor in the number of pokemon belonging to an opposing type and factor this into our overall interest in deciding which type has the highest overall advantage. We're going to query a different endpoint in the PokeAPI for this, `Pokemon` instead of `types`. We're going to maintain a dataframe that's incrementing the counts for each type. Note that we're going to count a Pokemon that has more than one type as one in each category for this purpose.  
  
```{r generatedis, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, cache = TRUE}
generate_distribution <- function(){
  counts <- rep(0, 18)
  types <- c("normal", "fire", "water", "electric", "grass", "ice", "fighting", "poison", "ground",          "flying", "psychic", "bug", "rock", "ghost", "dragon", "dark", "steel", "fairy")
  
  for (i in  1:905)
  {
    poke <- fromJSON(paste0("https://pokeapi.co/api/v2/pokemon/",i,"/"))
    for (t in poke$types$type$name)
    {
      counts[grep(t, types)] = counts[grep(t, types)] + 1
    }
  }
  
  return(counts)
}

# Taking a look at our raw counts
poke_counts <- generate_distribution()
poke_counts
```
  
# Exploratory Data Analysis
  
## Type Distribution  
We'll start our analysis here with taking a look at the distribution of Pokemon by type using the data we generated in our last function and make a histogram so we can see some of the discrepancies between the counts of the various types.  
  
```{r barchart, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, cache = TRUE}
types <- c("normal", "fire", "water", "electric", "grass", "ice", "fighting", "poison", "ground",          "flying", "psychic", "bug", "rock", "ghost", "dragon", "dark", "steel", "fairy")
df <- data.frame(poke_counts, row.names = types)
ggplot(df, aes(types, poke_counts)) + geom_col(aes(fill = types)) + ylab("# of Pokemon") + xlab("Pokemon Type") + ggtitle("Pokemon Count by Type")
```
  
## Combining Type Distribution with Matchup Matrix
Given the distribution of all the Pokemon types above, it would be interesting to see how this affects our calculated matchup advantages. The general premise would be that if a certain type has an advantage against a type that's more common, that advtange would be worth even more since it would theoretically come up more often in a matchup.  
  
We'll start with a scatter plot comparing the sums of all the advantages we calculated against all other types vs the number of that type of Pokemon that exist to get a general sense of what we're looking at here. As we can see below, we're kind of all over the place!  
  
```{r scatter, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, cache = TRUE}
sums <- rep(0, 18)
for(t in types)
{
  sums[grep(t, types)] = sum(matchups[t,])
}

df2 <- data.frame(cbind(df, sums))

ggplot(df2, aes(x = sums, y = poke_counts)) + geom_point(color = "#FC4E07") + xlab("Sum of Advantage Calculations") + ylab("# of Pokemon") + ggtitle("Type Advantage Sum vs # of Pokemon")
```

An interesting combination of these things could be to multiply the advantage calculation by the relative frequency of each type so that we get a sense of which types have the advantage against other types that have the most members. We'll make a new matrix for this that features the calculations and show the results in a barchart.  
  
```{r heatmap, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, cache = TRUE}
relative <- rep(0, 18)
for(t in types)
{
  relative[grep(t, types)] = df[t,]/1354
}
df3 <- cbind(df2, relative)
df3$final <- df3$relative * df3$sum

ggplot(df3, aes(x=types, y=final)) + geom_col(aes(fill = types)) + ylab("Overall Matchup Advantage") + ggtitle("Type Analysis")
```
