---
title: "Project 1 - API Vignette"
author: "Tommy King"
date: "`r Sys.Date()`"
output:
  github_document:
    html_preview: false
    toc: true
    toc_dept: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path = "../Images/")
```  
  
# Introduction  
In this vignette, we're going to be taking a look at the PokeAPI, writing some functions to interact with it, and doing some exploratory data analysis on the information we're grabbing.  
  
# Requirements  
For the work done here, we will be using the `tidyverse` package for all of our data manipulation, the `httr` package for making https requests to the API and `jsonlite` for interacting with json files. The code chunk for loading in these packages is included below  
  
```{r packages, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(httr)
library(jsonlite)
```  
  
# Functions Written for (Gotta) Catching All of the Data  
  
## Damage Type Functions  
In this section of functions, we will make some tools for querying the Pokemon type data so we can do some analysis on the balance between various types and their damage effectiveness vs. other types.
  
### It's super effective!  
This first function we're going to write is going to take a Pokemon type and utilize the `types` endpoint in the PokeAPI to return a list of types that the given type is super effective against. This will help us out later on in our analysis when we want to look at how many types each type is super effective against. We take the desired type as an argument here, grab the JSON from the API and drill down into the attribute `double_damage_to` which will give us a vector of the super effective types.  
  
```{r supereffective, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
superE <- function(type){
  type_info <- fromJSON(paste0("https://pokeapi.co/api/v2/type/",type,"/"))
  
  return(type_info$damage_relations$double_damage_to$name)
}
```
  
### It's not very effective...  
The next function we'll add is very similar to the previous one, but instead of getting the types our given type is super effective against, we'll grab the types that our given one only does half damage against.  
  
```{r notveryeffective, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
notVery <- function(type){
  type_info <- fromJSON(paste0("https://pokeapi.co/api/v2/type/",type,"/"))
  
  return(type_info$damage_relations$half_damage_to$name)
}
```
  
### It doesn't affect...
In the same vein here, this function will grab any types that the given type don't do any damage aginst. We'll eventually combine this with the previous function to compile any "bad" outcomes, ones where the given type isn't doing as much damage as it should.  
  
```{r doesntaffect, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
noEffect <- function(type){
  type_info <- fromJSON(paste0("https://pokeapi.co/api/v2/type/",type,"/"))
  
  return(type_info$damage_relations$no_damage_to$name)
}
```